---
title: "Untitled"
author: "Katherine Gong"
date: "2023-05-14"
output: pdf_document
---

Chapter 2 Statistical Learning

Lab: Introduction to R
```{r}
library(ISLR)
head(Auto)
```
```{r}
Auto_new <- na.omit(Auto)
dim(Auto_new)
```

```{r}
names(Auto_new)
```
```{r}
plot(Auto$cylinders, Auto$mpg)
```
The cylinders variable is stored as a numeric vector, so R has treated it as quantitative. However, since there are only a small number of possible values for cylinders, one may prefer to treat it as a qualitative variable. The as.factor() function converts quantitative variables into qualitative variables.
In R, factors are used to work with categorical variables, variables that have a fixed and known set of possible values. They are also useful when you want to display character vectors in a non-alphabetical order.
```{r}
# varwidth is a logical value, set as true to draw width of the box proportionate to the sample size
Auto_new$cyliners <- as.factor(Auto_new$cylinders)
plot(Auto_new$cylinders, Auto_new$mpg)
plot(Auto_new$cylinders, Auto_new$mpg, col = "red")
boxplot(Auto_new$cylinders, Auto_new$mpg, col = "red", varwidth = T)
boxplot(Auto_new$cylinders, Auto_new$mpg, col = "red", varwidth = T, horizontal = T)
boxplot(Auto_new$cylinders, Auto_new$mpg, col = "red", varwidth = T, xlab = "cylinders", ylab = "MPG")
```
The hist() function can be used to plot a histogram. Note that col = 2 has the same effect as col = "red"
```{r}
hist(Auto_new$mpg)
hist(Auto_new$mpg, col = 2)
hist(Auto_new$mpg, col = 2, breaks = 15) #breaks is used to mention the width of each bar
```
The pairs() function creates a scatterplot matrix, i.e. a scatterplot for every pair of variables. We can also produce scatterplots for just a subset of the variables.
```{r}
pairs(Auto_new)
pairs(~mpg + displacement + horsepower + weight + acceleration, data = Auto_new)
```
```{r}
summary(Auto_new)
```
```{r}
summary(Auto_new$mpg)
```

Graphics
The function seq() can be used to create a sequence of numbers. For instance, seq(a, b) makes a vector of integers between a and b. There are many other options: for instance, seq(0, 1,length = 10) makes a sequence of 10 numbers that are equally spaced between 0 and 1. Typing 3:11 is a shorthand for seq(3, 11) for integer arguments.
```{r}
x <- seq(1,10)
x
```
```{r}
x <- 1:10
x
```
```{r}
x <- seq(-pi, pi, length = 50)
x
```
The contour() function produces a contour plot in oreder to represent three-dimensional data; it is like a topographical map. It takes three arguments.
1. A vector of the x values (the first dimension)
2. A vector of the y values (the second dimension), and
3. A matrix whose elements correspond to the z value (the third dimension) for each pair of (x,y) coorinates.

As with the plot() function, there are many other inputs that can be used to fine-tune the output of the contour() function. 
```{r}
y <- x
f <- outer(x, y, function(x, y) cos(y) / (1+x^2))
contour(x,y,f)
contour(x, y, f, nlevels = 45, add = T)
```
```{r}
fa <- (f-t(f)) / 2
contour(x, y, fa, nlevels = 15)
```
The image() function works the same way as contour(), except that it produces a color-coded plot whose colors depend on the z value. This is know as a heatmap, and it somtimes used to plot temperature in weather forecasts. Alternatively, persp() can be used to produce a three-dimensional plot. The arguments theta and phi control the angles at which the plot is viewed.
```{r}
image(x, y, fa)
persp(x, y, fa)
persp(x, y, fa, theta = 30, phi = 20)
persp(x, y, fa, theta = 30, phi = 70)
persp(x, y, fa, theta = 30, phi = 40)
#angles defining the viewing direction. theta gives the azimuthal direction and phi the colatitude.
```

